Parallel multi-objective optimization using co-evolutionary islands with overlapping search space

### 2016-12-18 hpmoon-coor.tex
New title suggestion
Please @fergunet check it out.
Improving figure layout

\title{Parallel multi-objective optimization using co-evolutionary islands and search space overlapping}

@@

\begin{abstract}
Multi-objective optimization methods have gained interest in the last years because the performance they can achieve solving real-world problems. Many of these problems usually require a high-dimensional decision space, and multi-objective Evolutionary Algorithms have been successfully used in the past. Due to the large computational cost these algorithms require, different parallelization and distribution methods have been proposed, with different parallel and distributed computer architectures to cope with them. One of the most extended methods are the distributed co-evolutionary algorithms, where each processor deals only with a part of the decision space, that is, the individuals belonging to the same sub-population (island) explore the same subset of decision variables. These subsets are usually disjoint sets.

In our previous work we demonstrated that extending the length of these subsets, and therefore, sharing some search space between islands, implies an improvement in the results. In this paper, a new method to automatically adapt the size of the overlapping, or shared, sections of the individuals has been compared to other collaborative evolutionary techniques. Different number of islands, chromosome sizes and problems have been used. The analysis of the obtained experimental results, by using different metrics, shows that coevolutionary approaches can provide statistically significant improvements with respect to the base algorithm, being the method to automatically adapt the overlapping size the one to obtain the best performance during the same execution time. Results also show that the relation of the number of islands (subpopulations) to the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to distribute the decision variables.
\end{abstract}

@@

Different approaches have been used to parallelize Evolutionary Algorithms (EAs), as each individual can be considered as an independent unit \citep{Alba13parallel}. Classic methods, such as the global parallel EAs (Master-Slave), or the spatially structured algorithms (Island model or Cellular EAs) have been applied successfully in the past \citep{Folino03cellular,Alba02Parallelism}.  In the case of Multi-Objective Evolutionary Algorithms (MOEAs), these approaches \citep{Luna15Survey} need to deal with a whole solution set called Pareto Front (PFs). This imply to use different distribution and sharing mechanisms, as there exist a tension between the speedup achievable from parallelization and the need to globally recombine the results to accurately identify the PF \citep{Branke04Parallelizingcone}.

@@

In our preliminary work in \citep{Garcia16hpmoon} we tried to demonstrate that, instead focus every island in a disjoint subset of the chromosome, the usage of overlapped (shared) sections of the chromosome in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time when the number of islands increases. We discovered that the most adequate overlapping size may vary depending on a relation between the number of islands and the individual size. Moreover, the experiments were conducted in a mono-processor and synchronous island model. This motivates us to continue this research line, but this time, proposing a new method to automatically adapt the size of the overlapped sections, but also, performing all the experiments in a real cluster with up to 128 nodes, with extra number of islands.

The hypothesis of this paper is that using overlapped sections of the chromosome, calculated depending the number of islands to use, in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time with respect to the baseline, disjoint and partly overlapped methods. To demonstrate this, different overlapping islands schemes have been compared: a baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a coevolutionary disjoint section approach version of it, a fixed-size overlapped version and our new adaptive method. Different benchmarks problems and different large number of islands have been used. Results show that this new technique can improve different quality indicators in the same amount of time when the problem size is large.

@@

\subsection{Coevolution with disjoint islands (D)} 
In this approach, each individual of size $L$ is split into $P$ chunks of size $L/P$. Every island $p$ only performs crossover and mutation on the $p_{th}$ part of the individuals. Figure \ref{fig:disjoint} describes this approach.

\begin{figure*}
\centering
\includegraphics[width=12cm]{islandDisjoint.jpg}
\caption{Disjoint algorithm: every island $p$ only modifies the $p_{th}$ components (in grey) of the individuals.}
\label{fig:disjoint}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=12cm]{islandNoDisjoint.jpg}
\caption{Overlapping algorithm: every island $p$ modifies the  $p+1$,
  $p_{th}$ and $p-1$  components (in grey) of the individuals.}
  \label{fig:overlapping}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=12cm]{islandAdaptive.jpg}
\caption{Adaptive overlapping algorithm: every island $p$ modifies the  $p+c$,
  $p_{th}$ and $p-c$  components (in grey) of the individuals. $c$ is calculated depending on the number of islands. In this case, $c=2$.}
  \label{fig:adaptive}
\end{figure*}
% --------------------------------------------------------------

>>>>

\title{Parallel multi-objective optimization using co-evolutionary
  islands with overlapping search space}
% New title suggestion

@@

\begin{abstract}
  % First, state the problem!!!! - JJ
Multi-objective optimization methods have gained interest in the last
years because of the performance they can achieve solving real-world
problems. Many of these problems usually require a high-dimensional
decision space, and multi-objective Evolutionary Algorithms have been
successfully used in the past. Due to the large computational cost
these algorithms require, different parallelization and distribution
methods have been proposed, with different parallel and distributed
computer architectures to cope with them. One of the most extended
methods are the distributed co-evolutionary algorithms, where each
processor deals only with a part of the decision space, that is, the
individuals belonging to the same sub-population (island) explore the
same subset of decision variables. These subsets are usually disjoint
sets. We have previously proved that extending the length of these
subsets, and therefore, sharing some search space between islands,
implies an improvement in the results. In this paper, a new method to
automatically adapt the size of the overlapping, or shared, sections
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that coevolutionary
approaches can provide statistically significant improvements with
respect to the base algorithm, being the method to automatically adapt
the overlapping size the one to obtain the best performance during the
same execution time. Results also show that the relation of the number
of islands (subpopulations) to the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

%Introduction should start in the same way as the abstract. And
%starting with something that is not really applicable here, talking
%about general parallel evolutionary algorithms, is even worse - JJ
Different approaches have been used to parallelize Evolutionary
Algorithms (EAs), as each individual can be considered as an
independent unit \citep{Alba13parallel}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
In the case of Multi-Objective Evolutionary Algorithms (MOEAs), these
approaches \citep{Luna15Survey} need to deal with a whole solution set
called Pareto Front (PFs). This imply to use different distribution
and sharing mechanisms, as there exist a tension between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}.

@@

In our preliminary work \citep{Garcia16hpmoon} we tried to demonstrate
% tried to or actually did it? What did we prove? 
that,  instead of making every island focus in a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome in a coevolutionary multi-objective island-based algorithm
can improve the quality of the solutions in the same computing time
when the number of islands increases. We discovered that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. Moreover, the experiments
were conducted in a mono-processor and synchronous island model. This
motivates us to continue this research line; in this paper we propose
a new method that automatically adapts the size of the overlapped
sections. In order to test the validity of the method we have made
tests in a real cluster
with up to 128 nodes, with extra number of islands. % We have to
                                % enphatize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ

The hypothesis of this paper is that using overlapped sections of the
chromosome, calculated depending the number of islands to use, in a
coevolutionary multi-objective island-based algorithm can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. %That was the
                                %hypothesis of the previous
                                %paper. This paper has a new
                                %hypothesis, right? - JJ
In order to prove this claim, our new adaptive method has been
compared different overlapping islands schemes have been compared: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
coevolutionary disjoint section approach version of it and a
fixed-size overlapped version. %What was done in the previous paper?
                               %Are we re-using the results in that
                               %paper? - JJ
Different benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.  % Shouldn't this start with "We are trying to
                        % prove..." - JJ

@@

\subsection{Coevolution with disjoint islands (D)} 
In this approach, each individual of size $L$ is split into $P$ chunks of size $L/P$. Every island $p$ only performs crossover and mutation on the $p_{th}$ part of the individuals. Figure \ref{fig:disjoint} describes this approach.  
%
\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandDisjoint.jpg}
\caption{Disjoint algorithm: every island $p$ only modifies the $p_{th}$ components (in grey) of the individuals.}
\label{fig:disjoint}
\end{figure*}

@@

\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandNoDisjoint.jpg}
\caption{Overlapping algorithm: every island $p$ modifies the  $p+1$,
  $p_{th}$ and $p-1$  components (in grey) of the individuals.}
  \label{fig:overlapping}
\end{figure*}

@@

\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandAdaptive.jpg}
\caption{Adaptive overlapping algorithm: every island $p$ modifies the  $p+c$,
  $p_{th}$ and $p-c$  components (in grey) of the individuals. $c$ is calculated depending on the number of islands. In this case, $c=2$.}
  \label{fig:adaptive}
\end{figure*}
% --------------------------------------------------------------
